A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-07-23 survey.tex
Some revisions due to reviewers's comments

\section{Introduction}\label{sec:intro}
\IEEEPARstart{T}{he} field of real-time strategy (RTS) game AI has advanced significantly since Michael Buro's call for research in this area \cite{Buro03rts}. Specially, competitions like the ``AIIDE StarCraft AI Competition'' have motivated the exploration of many AI approaches in the context of RTS AI. We will list and classify these approaches, explain their 
strengths and weaknesses and conclude on what is left to achieve human-level 
RTS AI.

@@

The reader can find a  good presentation of task decomposition for AIs
playing RTS in~\cite{weber2011acs}. Although the previous task decomposition is common, a significant challenge is on designing architectures so that the individual AI techniques that address each of those tasks can communicate and effectively work together, resolving conflicts, prioritizing resources between them, etc. Section \ref{sec:bot} provides an overview of the task decompositions that state-of-the-art bots use.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{figures/levels_abstraction.pdf}
    \caption{RTS AI levels of abstraction and theirs properties: uncertainty (from intentions and from partial observation) is going higher as the abstraction levels are raising. Timings on the right correspond to an estimate of the duration of a behavior switch in StarCraft. Spatial and temporal reasoning are indicated for part for which greedy solutions are really not efficient.}
    \label{fig:levels-abstraction}
\end{figure}

@@

% PLANNING:
Approaches using planning techniques have also been explored in the literature. For example Onta\~{n}\'{o}n et al. \cite{CBR_Planning} explored the use of real-time case-based planning (CBP) in the domain of Wargus (a Warcraft II clone). In their work, they used human demonstration to learn plans, which are then composed at run-time in order to form full-fledges strategies to play the game. In \cite{PlanRetrieval} they improve over  their previous CBP approach by using situation assessment for improving the quality and speed of plan retrieval. Hierarchical Task-Network (HTN) planning has also been explored with some success in the context of simpler first-person shooter games \cite{HTNPlanning}. Planning approaches offer more adaptivity of the AI strategy compared to hard-coded approaches. However, the real-time constraints of RTS games limit the planning approaches that can be applied, being HTN and case-based planning being the only ones explored so far. Moreover, none of these approaches addresses any timing or scheduling issues, which are key in RTS games. On notable exception is the work of Churchill and Buro \cite{churchill2011build}, who used planning in order to construct its economic build-orders taking into account timing constraints of the different actions.

@@

% SCOUTING:
One final consideration concerning strategy is that RTS games are typically partially observable. Games like StarCraft implement the ``fog-of-war'' idea, which basically means that a player can only see the areas of the map close to her own units. Areas of the map away from the field of view of individual units are not observable. Players need to scout in order to obtain information about the opponent's strategy. The size of the state space in StarCraft prevents solutions based on POMDPs from being directly applicable, and very few of the previous approaches deal with this problem. Most research in RTS game AI assumes perfect information all the time, this mean that they only consider the information that they can see to make decisions. In the case of commercial games, most AI implementations cheat, since they have perfect information, while the human player does not. In order to make the human player believe the AI of these games does not cheat, sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft and StarCraft game series. A notable exception is the work of Weber et al. \cite{WeberAIIDE11}, who used a particle model with a linear trajectory update to track opponent units under fog-of-war in StarCraft. They also produced tactical goals through reactive planning and goal-driven autonomy \cite{WeberCig10,Weber10}, finding the more relevant goal(s) to spawn in unforeseen situations.

@@

% MACHINE LEARNING:
Concerning tactical decision making, many different approaches have been explored such as machine learning or game tree search. 
Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in first person shooter games (FPS) units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. Finally, \cite{SynnaeveTactics} combined space abstraction into regions from \cite{Perkins10} and tactical-decision making by assigning scores (economical, defenses, etc.) to regions and looking for their correspondences to tactical moves (attacks) in pro-gamers replays.
% GAME TREE SEARCH:
Game tree search techniques have also been explored for tactical decision making. \cite{Chung05} applied Monte-Carlo planning to a capture-the-flag version of Open RTS. Balla and Fern \cite{UCT} applied the UCT algorithm (a Monte Carlo Tree Search algorithm) to tactical assault planning in Wargus. To make game tree search applicable at this level, abstract game state representations are used in order to reduce the complexity. Also, abstractions, or simplifications about the set of possible actions to execute in a given game state need to be used.

@@

% POTENTIAL FIELDS:
Potential fields and influence maps have been found to be a useful technique for reactive decision making. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire \cite{uriarte2012kiting}, or staying at maximum shooting distance \cite{Hagelback09}. Potential fields have also been combined with A* path-finding to avoid local traps \cite{Hagelback12}. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multiagent potential fields based bot able to deal with fog-of-war in the Tankbattle game. Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial reasoning in RTS games. Danielsiek et al. \cite{Danielsiek_2008} used influence maps to achieve intelligent squad movement to flank the opponent in a RTS game. Despite their success, a drawback for potential field-based techniques is the large number of parameters that has to be tuned in order to achieve the desired behavior. Approaches for automatically learning such parameters have been explored, for example, using reinforcement learning \cite{Liu_2008}, or self-organizing-maps (SOM) \cite{teamCompositionRTS}. We would like to note that potential fields are a reactive control technique, and as such, they do not perform any form of lookahead. As a consequence, these techniques are prone to make units stuck in local maxima.

>>>>

\section{Introduction}\label{sec:intro}
\IEEEPARstart{T}{he} field of real-time strategy (RTS) game AI has advanced significantly since Michael Buro's call for research in this area \cite{Buro03rts}. Specially, competitions like the ``ORTS RTS Game AI Competition'' (held from 2006 to 2009), and the ``AIIDE StarCraft AI Competition'' (held since 2010) have motivated the exploration of many AI approaches in the context of RTS AI. We will list and classify these approaches, explain their 
strengths and weaknesses and conclude on what is left to achieve human-level 
RTS AI.

@@

The reader can find a  good presentation of task decomposition for AIs
playing RTS in~\cite{weber2011acs}. Although the previous task decomposition is common, a significant challenge is on designing architectures so that the individual AI techniques that address each of those tasks can communicate and effectively work together, resolving conflicts, prioritizing resources between them, etc. Section \ref{sec:bot} provides an overview of the task decompositions that state-of-the-art bots use. {\color{blue} Moreover, we would like to point out that the task decomposition above is not the only possible approach. Some systems, such as IMAI \cite{miles2007co}, divide gameplay into much smaller tasks, which are then assigned resources depending on the expected benefits of achieving each task.}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{figures/levels_abstraction.pdf}
    \caption{RTS AI levels of abstraction and theirs properties: uncertainty (coming from partial observation and from not knowing the intentions of the opponent) is higher for higher abstraction levels. Timings on the right correspond to an estimate of the duration of a behavior switch in StarCraft. Spatial and temporal reasoning are indicated for the levels at which greedy solutions are not enough.}
    \label{fig:levels-abstraction}
\end{figure}

@@

% PLANNING:
Approaches using planning techniques have also been explored in the literature. For example Onta\~{n}\'{o}n et al. \cite{CBR_Planning} explored the use of real-time case-based planning (CBP) in the domain of Wargus (a Warcraft II clone). In their work, they used human demonstration to learn plans, which are then composed at run-time in order to form full-fledges strategies to play the game. In \cite{PlanRetrieval} they improve over  their previous CBP approach by using situation assessment for improving the quality and speed of plan retrieval. Hierarchical Task-Network (HTN) planning has also been explored with some success in the context of simpler first-person shooter games \cite{HTNPlanning}. Planning approaches offer more adaptivity of the AI strategy compared to hard-coded approaches. However, the real-time constraints of RTS games limit the planning approaches that can be applied, HTN and case-based planning being the only ones explored so far. Moreover, none of these approaches addresses any timing or scheduling issues, which are key in RTS games. On notable exception is the work of Churchill and Buro \cite{churchill2011build}, who used planning in order to construct its economic build-orders taking into account timing constraints of the different actions.

@@

% SCOUTING:
One final consideration concerning strategy is that RTS games are typically partially observable. Games like StarCraft implement the ``fog-of-war'' idea, which basically means that a player can only see the areas of the map close to her own units. Areas of the map away from the field of view of individual units are not observable. Players need to scout in order to obtain information about the opponent's strategy. The size of the state space in StarCraft prevents solutions based on POMDPs from being directly applicable, and very few of the previous approaches deal with this problem. {\color{blue} Much work in RTS game AI assumes perfect information all the time. For example, in the case of commercial games, most AI implementations cheat, since the AI can see the complete game map at all times, while the human player does not. In order to make the human player believe the AI of these games does not cheat, sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft and StarCraft game series. Even if the StarCraft AI competition enforces fog-of-war, which means that bots are forced to work under partial information, little published research exists on this topic.} A notable exception is the work of Weber et al. \cite{WeberAIIDE11}, who used a particle model with a linear trajectory update to track opponent units under fog-of-war in StarCraft. They also produced tactical goals through reactive planning and goal-driven autonomy \cite{WeberCig10,Weber10}, finding the more relevant goal(s) to spawn in unforeseen situations.

@@

% MACHINE LEARNING:
Concerning tactical decision making, many different approaches have been explored such as machine learning or game tree search. 
Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in first person shooter games (FPS) units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. \cite{SynnaeveTactics} combined space abstraction into regions from \cite{Perkins10} and tactical-decision making by assigning scores (economical, defenses, etc.) to regions and looking for their correspondences to tactical moves (attacks) in pro-gamers replays.
{\color{blue} Finally, Christopher Miles \cite{miles2006co} created the idea of {\em IMTrees}, a tree where each leaf node is an influence map, and each intermediate node is a combination operation (sum, multiplication); Miles used evolutionary algorithms to learn IMTrees for each strategic decision in the game involving spatial reasoning by combining a set of basic influence maps.}

% GAME TREE SEARCH:
Game tree search techniques have also been explored for tactical decision making. \cite{Chung05} applied Monte-Carlo planning to a capture-the-flag version of Open RTS. Balla and Fern \cite{UCT} applied the UCT algorithm (a Monte Carlo Tree Search algorithm) to tactical assault planning in Wargus. To make game tree search applicable at this level, abstract game state representations are used in order to reduce the complexity. Also, abstractions, or simplifications about the set of possible actions to execute in a given game state need to be used.

@@

% POTENTIAL FIELDS:
Potential fields and influence maps have been found to be useful techniques for reactive decision making. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire \cite{uriarte2012kiting}, or staying at maximum shooting distance \cite{Hagelback09}. Potential fields have also been combined with A* path-finding to avoid local traps \cite{Hagelback12}. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multiagent potential fields based bot able to deal with fog-of-war in the Tankbattle game. Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial reasoning in RTS games. Danielsiek et al. \cite{Danielsiek_2008} used influence maps to achieve intelligent squad movement to flank the opponent in a RTS game. Despite their success, a drawback for potential field-based techniques is the large number of parameters that has to be tuned in order to achieve the desired behavior. Approaches for automatically learning such parameters have been explored, for example, using reinforcement learning \cite{Liu_2008}, or self-organizing-maps (SOM) \cite{teamCompositionRTS}. We would like to note that potential fields are a reactive control technique, and as such, they do not perform any form of lookahead. As a consequence, these techniques are prone to make units stuck in local maxima.
